---
title: "R for data science"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Preparation

```{r}
# install.packages("tidyverse")
# install.packages(c("nycflights13", "gapminder", "Lahman"))
```


# Part 1. Explore 

```{r}
library(tidyverse)
```


## 1. Data Visualization with ggplot2

```{r}
library(tidyverse)
mpg
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap( ~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, color = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv), show.legend = FALSE)
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_point() + geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_point(mapping = aes(color = class)) + geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_point(mapping = aes(color = class)) + geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = diamonds) + stat_count(mapping = aes(x = cut))
```

```{r}
demo <- tribble(~a, ~b, 
                "bar_1", 20,
                "bar_2", 30,
                "bar_3", 40)
ggplot(data = demo) + geom_bar(mapping = aes(x = a, y = b), stat = 'identity')
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

```{r}
ggplot(data = diamonds) + stat_summary(mapping = aes(x = cut, y = depth), 
                                       fun.ymin = min, fun.ymax = max,
                                       fun.y = median)
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, color = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, color = clarity)) + 
  geom_bar(fill = NA, position = "identity")
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + geom_boxplot()
```

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + geom_boxplot() + coord_flip()
```

```{r}
nz <- map_data("nz")
ggplot(nz, aes(long, lat, group = group)) + 
  geom_polygon(fill = "white", color = "black")
```

```{r}
ggplot(nz, aes(long, lat, group = group)) + 
  geom_polygon(fill = "white", color = "black") + 
  coord_quickmap()
```

```{r}
bar <- ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut),
                                          show.legend = FALSE,
                                          width = 1) + theme(aspect.ratio = 1) + labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```


## 2. Workflow: Basic 

```{r}
seq(1,10)
```

```{r}
(y <- seq(1,10,length.out = 5))
```

## 3. Data transformation with dplyr 

```{r}
library(nycflights13)
library(tidyverse)
```

```{r}
flights
```

```{r}
class(flights)
```

```{r}
filter(flights, month == 1, day == 1)
```

```{r}
jan1 <- filter(flights, month == 1, day == 1)
(dec25 <- filter(flights, month == 12, day == 25))
```

```{r}
filter(flights, month == 11 | month == 12)
```

```{r}
nov_dec <- filter(flights, month %in% c(11,12))
```

```{r}
filter(flights, !(arr_delay < 120 | dep_delay > 120))
```

```{r}
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

```{r}
x = NA
is.na(x)
```

```{r}
# filter only includes rows where the condition is TRUE, it excludes both FALSE and NA values. 
df <- tibble(x = c(1,NA,3))
filter(df, x > 1)
```

```{r}
filter(df, is.na(x) | x > 1)
```

```{r}
arrange(flights, year, month, day)
```

```{r}
arrange(flights, desc(arr_delay))
```

```{r}
# missing values are always sorted at the end 
df <- tibble(x = c(5,2,NA))
arrange(df, x)
```

```{r}
select(flights, year, month, day)
```

```{r}
select(flights, year:day)
```

```{r}
select(flights, -(year:day))
```

```{r}
rename(flights, tail_num = tailnum)
```

```{r}
select(flights, time_hour, air_time, everything())
```

```{r}
flights_sml <- select(flights, 
                      year:day,
                      ends_with("delay"),
                      distance, 
                      air_time)
mutate(flights_sml,
       gain = arr_delay - dep_delay,
       speed = distance / air_time * 60)
```

```{r}
mutate(flights_sml,
       gain = arr_delay - dep_delay,
       hours = air_time / 60,
       gain_per_hour = gain / hours)
```

```{r}
transmute(flights, 
          gain = arr_delay - dep_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hours)
```

```{r}
transmute(flights,
          dep_time, 
          hour = dep_time %/% 100,
          minute = dep_time %% 100)
```



```{r}
x <- 1:10
lag(x)
lead(x)
```

```{r}
cumsum(x)
cummean(x)
```

```{r}
v <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_day <- group_by(flights, year, month, day)
summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_dest <- group_by(flights, dest)
delay <- summarize(by_dest,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE))
delay <- filter(delay, count > 20, dest != "HNL")
```

```{r}
ggplot(data = delay, mapping = aes(x = dist, y = delay)) + 
  geom_point(aes(size = count), alpha = 1/3) + 
  geom_smooth(se = FALSE)
```

```{r}
delays <- flights %>% 
          group_by(dest) %>%
          summarize(count = n(),
                    dist = mean(distance, na.rm = TRUE),
                    delay = mean(arr_delay, na.rm = TRUE)) %>% 
          filter(count > 20, dest != "HNL")
```

```{r}
?ggvis
```

```{r}
flights %>% group_by(year, month, day) %>% 
            summarize(mean = mean(dep_delay))
```

```{r}
flights %>% 
        group_by(year, month, day) %>%
        summarize(mean = mean(dep_delay, na.rm = TRUE))
```

```{r}
not_cancelled <- flights %>% 
                         filter(!is.na(dep_delay), !is.na(arr_delay))
```

```{r}
not_cancelled %>% 
              group_by(year, month, day) %>% 
              summarize(mean = mean(dep_delay))
```

```{r}
delays <- not_cancelled %>% 
                        group_by(tailnum) %>% 
                        summarize(delay = mean(arr_delay))
ggplot(data = delays, mapping = aes(x = delay)) +
  geom_freqpoly(binwidth = 10)
```

```{r}
delays <- not_cancelled %>% 
                        group_by(tailnum) %>% 
                        summarize(delay = mean(arr_delay, na.rm = TRUE),
                                  n = n())
ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

```{r}
batting <- as_tibble(Lahman::Batting)

batters <- batting %>% 
  group_by(playerID) %>% 
  summarize(ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
            ab = sum(AB, na.rm = TRUE))

batters %>% 
  filter(ab > 100) %>% 
  ggplot(mapping = aes(x = ab, y = ba)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r}
batters %>% arrange(desc(ba))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(avg_delay1 = mean(arr_delay),
            avg_delay2 = mean(arr_delay[arr_delay > 0]))
```

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarize(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(first = min(dep_time),
            last = max(dep_time))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(first_dep = first(dep_time),
            last_dep = last(dep_time))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarize(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```

```{r}
not_cancelled %>% 
  count(dest)
```

```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```

```{r}
not_cancelled %>%
  group_by(year, month, day) %>% 
  summarize(n_early = sum(dep_time < 500))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(hour_perc = mean(arr_delay > 60))
```

```{r}
daily <- group_by(flights, year, month, day)
(per_day <- summarize(daily, flights = n()))
```

```{r}
(per_month <- summarize(per_day, flights = sum(flights)))
```

```{r}
(per_year <- summarize(per_month, flights = sum(flights)))
```

```{r}
daily %>% 
  ungroup() %>% 
  summarize(flights = n())
```

```{r}
flights_sml %>% 
  group_by(year, month, day) %>% 
  filter(rank(desc(arr_delay))<10)
```

```{r}
popular_dests <- flights %>%
  group_by(dest) %>%
  filter(n()>365)
popular_dests
```

```{r}
popular_dests %>%
  filter(arr_delay > 0) %>%
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>%
  select(year:day, dest, arr_delay, prop_delay)
  
```


## 4. Workflow: Scripts 

## 5. Exploratory data analysts

```{r}
library(tidyverse)
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

```{r}
diamonds %>% count(cut)
```

```{r}
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)
```

```{r}
diamonds %>% 
  count(cut_width(carat, 0.5))
```

```{r}
smaller <- diamonds %>% 
  filter(carat < 3)

ggplot(data = smaller, mapping = aes(x = carat)) + 
  geom_histogram(binwidth = 0.1)
```

```{r}
ggplot(data = smaller, mapping = aes(x = carat, color = cut)) + 
  geom_freqpoly(binwidth = 0.1)
```

```{r}
ggplot(data = smaller, mapping = aes(x = carat)) + geom_histogram(binwidth = 0.01)
```

```{r}
ggplot(data = faithful, mapping = aes(x = eruptions)) + geom_histogram(binwidth = 0.25)
```

```{r}
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5)
```

```{r}
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) + 
  coord_cartesian(ylim = c(0,50))
```

```{r}
unusuak <- diamonds %>% 
  filter(y < 3 | y > 20) %>% 
  arrange(y)
unusual
```

```{r}
diamonds2 <- diamonds %>% 
  filter(between(y, 3, 20))
```

```{r}
diamonds2 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y))
```

```{r}
ggplot(data = diamonds2, mapping = aes(x = x, y = y)) + 
  geom_point()
```

```{r}
ggplot(data = diamonds2, mapping = aes(x = x, y = y)) + 
  geom_point(na.rm = TRUE)
```

```{r}
nycflights13::flights %>% 
  mutate(cancelled = is.na(dep_time),
         sched_hour = sched_dep_time %/% 100,
         sched_min = sched_dep_time %% 100,
         sched_dep_time = sched_hour + sched_min / 60) %>% 
  ggplot(mapping = aes(sched_dep_time)) + 
  geom_freqpoly(mapping = aes(color = cancelled),
                binwidth = 1/4)
```

```{r}
ggplot(data = diamonds, mapping = aes(x = price)) + 
  geom_freqpoly(mapping = aes(color = cut), binwidth = 500)
```

```{r}
ggplot(diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(color = cut), binwidth = 500)
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = price)) + 
  geom_boxplot()
```

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
```

```{r}
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median),y = hwy))
```

```{r}
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy)) + 
  coord_flip()
```

```{r}
ggplot(data = diamonds) + 
  geom_count(mapping = aes(x = cut, y = color))
```

```{r}
diamonds %>% 
  count(color, cut)
```

```{r}
diamonds %>% 
  count(color, cut) %>% 
  ggplot(mapping = aes(x = color, y = cut)) + 
  geom_tile(mapping = aes(fill = n))
```

```{r}
ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price))
```

```{r}
ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price), alpha = 1/100)
```

```{r}
ggplot(data = smaller) + 
  geom_bin2d(mapping = aes(x = carat, y = price))
```


```{r}
ggplot(data = smaller) + 
  geom_hex(mapping = aes(x = carat, y = price))
```

```{r}
ggplot(data = smaller, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)))
```

```{r}
ggplot(data = smaller, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_number(carat, 20)))
```

```{r}
ggplot(data = faithful) + 
  geom_point(mapping = aes(x = eruptions, y = waiting))
```

```{r}
library(modelr)
mod <- lm(log(price) ~ log(carat), data = diamonds)

diamonds2 <- diamonds %>% 
  add_residuals(mod) %>% 
  mutate(resid = exp(resid))

ggplot(data = diamonds2) + 
  geom_point(mapping = aes(x = carat, y = resid))
```

```{r}
ggplot(data = diamonds2) + 
  geom_boxplot(mapping = aes(x = cut, y = resid))
```

```{r}
ggplot(data = faithful, mapping = aes(x = eruptions)) + 
  geom_freqpoly(bindwidth = 0.25)
```

```{r}
ggplot(faithful, aes(eruptions)) + 
  geom_freqpoly(binwidth = 0.25)
```

```{r}
diamonds %>% 
  count(cut, clarity) %>%
  ggplot(aes(clarity, cut, fill = n)) + 
  geom_tile()
```

## 6. Workflow: Projects 

# Part 2. Wrangle 

## 7. Tibbles with tibble

```{r}
library(tidyverse)
```

```{r}
as_tibble(iris)
```

```{r}
tibble(
  x = 1:5,
  y = 1,
  z = x ^ 2 + y
)
```

```{r}
tb <- tibble(
  `:)` = "smile",
  ` ` = "space",
  `2000` = "number"
)
tb
```

```{r}
tribble(
  ~x, ~y, ~z,
  "a",2,3.6,
  "b",1,8.5
)
```

```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```

```{r}
nycflights13::flights %>% print(n = 10, width = Inf)
```

```{r}
# options(tibble.print_max = n, tibble.print_min = m)
# options(tibble.width = Inf)
```

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)
```

```{r}
df$x
```

```{r}
df[["x"]]
```

```{r}
df["x"]
```

```{r}
df[[1]]
```

```{r}
df[1]
```

```{r}
df %>% .$x
```

```{r}
df %>% .[["x"]]
```

```{r}
class(as.data.frame(tb))
```

## 8. Data import with readr 

```{r}
library(tidyverse)
```

```{r}
# read_csv(), read_csv2(), read_tsv(), read_delim(), read_fwf(), read_log()
```

```{r}
heights <- read_csv("data/heights.csv")
```

```{r}
read_csv("a,b,c
         1,2,3
         4,5,6")
```

```{r}
read_csv("The first line of metadata
         The second line of metadata
         x,y,z
         1,2,3", skip = 2)
```

```{r}
read_csv("# A comment I want to skip
         x,y,z
         1,2,3", comment = "#")
```

```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```

```{r}
read_csv("1,2,3\n4,5,6", col_names = c("x","y","z"))
```

```{r}
read_csv("a,b,c\n1,2,.",na = ".")
```

```{r}
# compared to base R, faster, have progress bar, produce tibbles, don't convert character vectors to facors, use row names, reproducible
```

```{r}
str(parse_logical(c("TRUE","FALSE","NA")))
```

```{r}
str(parse_integer(c("1","2","3")))
```

```{r}
str(parse_date(c("2010-01-01","1979-10-14")))
```

```{r}
parse_integer(c("1","231",".","456"), na=".")
```

```{r}
x <- parse_integer(c("123","345","abc","123.45"))
```

```{r}
x
```

```{r}
problems(x)
```

```{r}
parse_double("1.23")
```

```{r}
parse_double("1,23", locale = locale(decimal_mark = ","))
```

```{r}
parse_number("$100")
parse_number("20%")
parse_number("It cost $123.45")
```

```{r}
parse_number("$123,456,789")
parse_number("123.456.789", locale = locale(grouping_mark = "."))
parse_number("123`456`789", locale = locale(grouping_mark = "`"))
```

```{r}
charToRaw("Hadley")
```

```{r}
x1 <- "El Ni\xf1o was particularly bad this year"
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"
```

```{r}
parse_character(x1, locale = locale(encoding = "Latin1"))
parse_character(x2, locale = locale(encoding = "Shift-JIS"))
```

```{r}
guess_encoding(charToRaw(x1))
```

```{r}
guess_encoding(charToRaw(x2))
```

```{r}
fruit <- c("apple", "banana")
parse_factor(c("apple","banana","bananana"), levels = fruit)
```

```{r}
parse_datetime("2010-10-01T2010")
parse_datetime("20101010")
```

```{r}
parse_date("2010-10-01")
```

```{r}
library(hms)
parse_time("01:10 am")
parse_time("20:10:01")
```

```{r}
parse_date("01/02/15","%m/%d/%y")
parse_date("01/02/15","%d/%m/%y")
parse_date("01/02/15","%y/%m/%d")
```

```{r}
parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
```

```{r}
guess_parser("2010-10-01")
guess_parser("15:01")
guess_parser(c("TRUE", "FALSE"))
guess_parser(c("1","5","9"))
guess_parser(c("12, 352, 561"))
str(parse_guess("2010-10-10"))
```

```{r}
challenge <- read_csv(readr_example("challenge.csv"))
```

```{r}
problems(challenge)
```

```{r}
challenge <- read_csv(readr_example("challenge.csv"),
                      col_types = cols(
                        x = col_integer(),
                        y = col_character()
                      ))
```

```{r}
challenge <- read_csv(
  readr_example("challenge.csv"),
  col_types = cols(
    x = col_double(),
    y = col_character()
  )
)
```

```{r}
tail(challenge)
```

```{r}
challenge <- read_csv(
  readr_example("challenge.csv"),
  col_type = cols(
    x = col_double(),
    y = col_date()
  )
)
```

```{r}
tail(challenge)
```

```{r}
challenge2 <- read_csv(
  readr_example("challenge.csv"),
  guess_max = 1001
)
challenge2
```

```{r}
challenge2 <- read_csv(readr_example("challenge.csv"),
                       col_types = cols(.default = col_character()))
```

```{r}
df <- tribble(
  ~x, ~y,
  "1", "1.21",
  "2", "2.32",
  "3", "4.56"
)
df
```

```{r}
type_convert(df)
```

```{r}
write_csv(challenge,"challenge.csv")
```

```{r}
challenge
write_csv(challenge, "challenge-2.csv")
read_csv("challenge-2.csv")
```

```{r}
write_rds(challenge, "challenge.rds")
read_rds("challenge.rds")
```

```{r}
library(feather)
write_feather(challenge, "challenge.feather")
read_feather("challenge.feather")
```


## 9. Tidy data with tidyr 

```{r}
table1
```

```{r}
table2
```

```{r}
table3
```

```{r}
table4a
```

```{r}
table4b
```

```{r}
table1 %>% 
  mutate(rate = cases / population * 10000)
```

```{r}
table1 %>%
  count(year, wt = cases)
```

```{r}
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), color = "grey50") + 
  geom_point(aes(color = country))
```

```{r}
table4a
```

```{r}
table4a %>%
  gather(`1999`,`2000`,key = "year", value = "cases")
```

```{r}
table4b
```

```{r}
table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
```

```{r}
tidy4a <- table4a %>%
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>%
  gather(`1999`, `2000`, key = "year", value = "population")
left_join(tidy4a, tidy4b)
```

```{r}
table2
```

```{r}
spread(table2, key = type, value = count)
```

```{r}
table3
```

```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"))
```

```{r}
table3 %>%
  separate(rate, into = c("cases", "population"), sep = "/")
```

```{r}
table3 %>%
  separate(
    rate, 
    into = c("cases","population"),
    convert = TRUE
  )
```

```{r}
table5
```

```{r}
table5 %>% 
  unite(new, century, year)
```

```{r}
table5 %>%
  unite(new, century, year, sep = "")
```

```{r}
stocks <- tibble(
  year = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr = c(1, 2, 3, 4, 2, 3, 4),
  return = c(1.88, 0.59, 0.35, NA, 0.92, 0.17, 2.66)
)
stocks
```

```{r}
stocks %>% 
  spread(year, return)
```

```{r}
stocks %>%
  spread(year, return) %>% 
  gather(year, return, `2015`, `2016`, na.rm = TRUE)
```

```{r}
stocks %>% 
  complete(year, qtr)
```

```{r}
treatment <- tribble(
  ~ person, ~ treatment, ~ response,
  "Derrick Whitmore", 1, 7,
  NA, 2, 10,
  NA, 3, 9,
  "Katherine Burke", 1, 4
)
treatment
```

```{r}
treatment %>%
  fill(person)
```

```{r}
who
```

```{r}
who1 <- who %>% 
  gather(new_sp_m014:newrel_f65, 
         key = "key", 
         value = "cases",
         na.rm = TRUE)
who1
```

```{r}
who1 %>%
  count(key)
```

```{r}
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
who2
```
```{r}
who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
who3
```

```{r}
who3 %>% 
  count(new)
```

```{r}
who4 <- who3 %>%
  select(-new, -iso2, -iso3)
```

```{r}
who5 <- who4 %>%
  separate(sexage, c("sex","age"),sep = 1)
who5
```

```{r}
who %>% 
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = stringr::str_replace(code, "newrel", "new_rel")) %>% 
  separate(code, c("new", "var", "sexage")) %>%
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```


## 10. Relational data with dplyr 

```{r}
library(tidyverse)
library(nycflights13)
```

```{r}
airlines
```

```{r}
airports
```

```{r}
planes
```

```{r}
weather
```

```{r}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)
```

```{r}
weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)
```

```{r}
flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)
```

```{r}
flights %>% 
  count(year, month, day, tailnum) %>% 
  filter(n > 1)
```

```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2
```

```{r}
flights2 %>% 
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")
```

```{r}
flights2 %>% 
  select(-origin, -dest) %>% 
  mutate(name = airlines$name[match(carrier, airlines$carrier)])
```

```{r}
x <- tribble(
  ~ key, ~val_X,
  1, "x1",
  2, "x2",
  3, "x3"
)
```

```{r}
y <- tribble(
  ~ key, ~ val_y,
  1, "y1",
  2, "y2",
  4, "y3"
)
```

```{r}
x %>% 
  inner_join(y, by = "key")
```

```{r}
x <- tribble(
  ~ key, ~val_x,
  1, "x1",
  2, "x2",
  2, "x3",
  1, "x4"
)
```

```{r}
y <- tribble(
  ~ key, ~val_y,
  1, "y1",
  2, "y2"
)
```

```{r}
left_join(x, y, by = "key")
```

```{r}
x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  2, "x3",
  3, "x4"
)
y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2",
  2, "y3",
  3, "y4"
)
left_join(x, y, by = "key")
```

```{r}
flights2 %>% 
  left_join(weather)
```

```{r}
flights2 %>% 
  left_join(planes, by = "tailnum")
```

```{r}
flights2 %>%
  left_join(airports, c("dest" = "faa"))
```

```{r}
flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```

```{r}
top_dest <- flights %>% 
  count(dest, sort = TRUE) %>% 
  head(10)
top_dest
```

```{r}
flights %>% 
  filter(dest %in% top_dest$dest)
```

```{r}
flights %>% 
  semi_join(top_dest)
```

```{r}
flights %>% 
  anti_join(planes, by = "tailnum") %>% 
  count(tailnum, sort = TRUE)
```

```{r}
airports %>% 
  count(alt, lon) %>% 
  filter(n>1)
```

```{r}
df1 <- tribble(
  ~x, ~y,
  1, 1,
  2, 1
)
df2 <- tribble(
  ~x, ~y,
  1, 1, 
  1, 2
)
```

```{r}
intersect(df1, df2)
```

```{r}
union(df1, df2)
```

```{r}
setdiff(df1, df2)
```

```{r}
setdiff(df2, df1)
```



## 11. Strings with stringr 

```{r}
library(tidyverse)
library(stringr)
```

```{r}
string1 <- "This is a string"
string2 <- 'To put a "quote" inside a string, use single quotes'
```

```{r}
double_quote <- "\"" # or '"'
single_quote <- '\'' # or "'"
```

```{r}
x <- c("\"", "\\")
x
writeLines(x)
```

```{r}
x <- "\u00b5"
x
```

```{r}
c("one", "two", "three")
```

```{r}
str_length(c("a","R for data science", NA))
```

```{r}
str_c("x", "y")
str_c("x","y","z")
```

```{r}
str_c("x","y",sep=", ")
```

```{r}
x <- c("abc", NA)
str_c("|-", str_replace_na(x), "-|")
```

```{r}
str_c("prefix-", c("a","b","c"),"-suffix")
```

```{r}
name <- "Hadley"
time_of_day <- "mornig"
birthday <- FALSE

str_c("Good ", time_of_day, " ", name, 
      if(birthday)" and HAPPY BIRTHDAY", ".")
```

```{r}
str_c(c("x","y","z"), collapse = ", ")
```

```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
str_sub(x, -3, -1)
```

```{r}
str_sub("a",1,5)
```

```{r}
str_sub(x,1,1) <- str_to_lower(str_sub(x,1,1))
x
```

```{r}
str_to_upper(c("i","ı"))
str_to_upper(c("i", "ı"), locale = "tr")
```

```{r}
x <- c("apple", "eggplant","banana")
str_sort(x, locale = "en")
str_sort(x, locale = "haw")
```

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```

```{r}
str_view(x, ".a.")
```

```{r}
dot <- "\\."
writeLines(dot)
str_view(c("abc","a.c","bef"),"a\\.c")
```

```{r}
x <- "a\\b"
writeLines(x)
str_view(x, "\\\\")
```

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
str_view(x, "a$")
```

```{r}
x <- c("apple pie","apple","apple cake")
str_view(x, "apple")
str_view(x, "^apple$")
```

```{r}
str_view(c("grey","gray"),"gr(e|a)y")
```

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
str_view(x, "CC+")
str_view(x, "C[LX]+")
```

```{r}
str_view(x, "C{2}")
str_view(x, "C{2,}")
str_view(x, "C{2,3}")
```

```{r}
str_view(x, "C{2,3}?")
str_view(x, "C[LX]+?")
```

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

```{r}
x <- c("apple","banana","pear")
str_detect(x, "e")
```

```{r}
sum(str_detect(words, "^t"))
mean(str_detect(words, "[aeiou]$"))
```

```{r}
no_vowels_1 <- !str_detect(words, "[aeiou]")
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

```{r}
words[str_detect(words,"x$")]
str_subset(words, "x$")
```

```{r}
df <- tibble(
  word = words,
  i = seq_along(word)
)
df
```


```{r}
df %>% 
  filter(str_detect(words, "x$"))
```


```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
mean(str_count(words, "[aeiou]"))
```

```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

```{r}
str_count("abababa","aba")
str_view_all("abababa","aba")
```

```{r}
length(sentences)
head(sentences)
```

```{r}
colors <- c("red", "orange", "yellow", "green", "blue", "purple")
color_match <- str_c(colors, collapse = "|")
color_match
```

```{r}
has_color <- str_subset(sentences, color_match)
matches <- str_extract(has_color, color_match)
head(matches)
```

```{r}
more <- sentences[str_count(sentences, color_match)>1]
str_view_all(more, color_match)
```

```{r}
str_extract(more, color_match)
```

```{r}
str_extract_all(more, color_match)
```

```{r}
str_extract_all(more, color_match, simplify = TRUE)
```

```{r}
x <- c("a","a b","a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

```{r}
noun <- "(a|the) ([^ ]+)"
has_noun <- sentences %>%
  str_subset(noun) %>% 
  head(10)
has_noun %>% 
  str_extract(noun)
```

```{r}
has_noun %>%
  str_match(noun)
```

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article","noun"),"(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

```{r}
sentences %>%
  head(5) %>%
  str_split(" ")
```

```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

```{r}
sentences %>%
  head(5) %>%
  str_split(" ", simplify = TRUE)
```

```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% 
  str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "This is a sentence. This is another sentence."
str_view_all(x, boundary("word"))
```

```{r}
str_split(x, " ")[[1]]
```

```{r}
str_split(x, boundary("word"))[[1]]
```

```{r}
str_view(fruit, "nana")
```

```{r}
str_view(fruit, regex("nana"))
```

```{r}
bananas <- c("banana", "Banana", "BANANA")
str_view(bananas, "banana")
```

```{r}
str_view(bananas, regex("banana", ignore_case = TRUE))
```

```{r}
x <- "Line 1\nLine 2\nLine 3"
str_extract_all(x, "^Line")[[1]]
str_extract_all(x, regex("^Line", multiline = TRUE))[[1]]
```

```{r}
phone <- regex("
    \\(? # optional opening parens
    (\\d{3}) # area code
    [)- ]? # optional closing parens, dash, or space
    (\\d{3}) # another three numbers
    [ -]? # optional space or dash
    (\\d{3}) # three more numbers
    ", comments = TRUE)
str_match("514-791-8141", phone)
```

```{r}
microbenchmark::microbenchmark(
  fixed = str_detect(sentences, fixed("the")),
  regex = str_detect(sentences, "the"),
  times = 20
)
```

```{r}
a1 <- "\u00e1"
a2 <- "a\u0301"
c(a1, a2)
a1 == a2
```

```{r}
str_detect(a1, fixed(a2))
str_detect(a1, coll(a2))
```

```{r}
i <- c("I", "İ", "i", "ı")
i

str_subset(i, coll("i", ignore_case = TRUE))

str_subset(
i,
coll("i", ignore_case = TRUE, locale = "tr")
)
```

```{r}
stringi::stri_locale_info()
```

```{r}
x <- "This is a sentence"
str_view_all(x, boundary("word"))
str_extract_all(x, boundary("word"))
```

```{r}
apropos("replace")
```

```{r}
head(dir(pattern = "\\.Rmd$"))
```



## 12. Factors with forcats 

```{r}
library(tidyverse)
library(forcats)
```

```{r}
x1 <- c("Dec","Apr","Jan","Mar")
x2 <- c("Dec","Apr","Jam","Mar")
sort(x1)
```

```{r}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun",
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
y1 <- factor(x1, levels = month_levels)
sort(y1)
```

```{r}
y2 <- factor(x2, levels = month_levels)
y2
```

```{r}
y2 <- parse_factor(x2, levels = month_levels)
```

```{r}
factor(x1)
```

```{r}
f1 <- factor(x1, levels = unique(x1))
f1 

f2 <- x1 %>%
  factor() %>%
  fct_inorder()
f2
levels(f2)
```

```{r}
gss_cat
```

```{r}
gss_cat %>% 
  count(race)
```

```{r}
ggplot(gss_cat, aes(race)) + geom_bar()
```

```{r}
ggplot(gss_cat, aes(race)) + geom_bar() + scale_x_discrete(drop = FALSE)
```

```{r}
relig <- gss_cat %>%
  group_by(relig) %>%
  summarize(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )
ggplot(relig, aes(tvhours, relig)) + geom_point()
```

```{r}
relig
```


```{r}
ggplot(relig, aes(tvhours, fct_reorder(relig, tvhours))) + geom_point()
```

```{r}
relig %>%
  mutate(relig = fct_reorder(relig, tvhours)) %>%
  ggplot(aes(tvhours, relig)) + geom_point()
```

```{r}
rincome <- gss_cat %>%
  group_by(rincome) %>%
  summarize(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )
ggplot(rincome, aes(age, fct_reorder(rincome, age))) + geom_point()
```

```{r}
ggplot(rincome, aes(age, fct_relevel(rincome, "Not applicable"))) + geom_point()
```

```{r}
by_age <- gss_cat %>% 
  filter(!is.na(age)) %>%
  group_by(age, marital) %>%
  count() %>%
  mutate(prop = n/sum(n))

ggplot(by_age, aes(age, prop, color = marital)) + 
  geom_line(na.rm = TRUE)

ggplot(by_age, aes(age, prop, color = fct_reorder2(marital, age, prop))) + 
  geom_line() + 
  labs(color = "marital")

```

```{r}
gss_cat %>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(marital)) + 
  geom_bar()
```

```{r}
gss_cat %>% count(partyid)
```

```{r}
gss_cat %>% 
  mutate(partyid = fct_recode(partyid,
"Republican, strong" = "Strong republican",
"Republican, weak" = "Not str republican",
"Independent, near rep" = "Ind,near rep",
"Independent, near dem" = "Ind,near dem",
"Democrat, weak" = "Not str democrat",
"Democrat, strong" = "Strong democrat"
                              )) %>%
  count(partyid)
```

```{r}
gss_cat %>%
  mutate(partyid = fct_recode(partyid,
"Republican, strong" = "Strong republican",
"Republican, weak" = "Not str republican",
"Independent, near rep" = "Ind,near rep",
"Independent, near dem" = "Ind,near dem",
"Democrat, weak" = "Not str democrat",
"Democrat, strong" = "Strong democrat",
"Other" = "No answer",
"Other" = "Don't know",
"Other" = "Other party"              
                              )) %>%
  count(partyid)
```

```{r}
gss_cat %>% 
  mutate(partyid = fct_collapse(partyid,
        other = c("No answer", "Don't know", "Other party"),
        rep = c("Strong republican", "Not str republican"),
        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
        dem = c("Not str democrat", "Strong democrat"))) %>%
  count(partyid)
```

```{r}
gss_cat %>% 
  mutate(relig = fct_lump(relig)) %>%
  count(relig)
```

```{r}
gss_cat %>% 
  mutate(relig = fct_lump(relig, n = 10)) %>%
  count(relig, sort = TRUE) %>%
  print(n = Inf)
```



## 13. Dates and times with lubridate 


```{r}
library(tidyverse)
library(lubridate)
library(nycflights13)
```

```{r}
today()
now()
```

```{r}
ymd("2017-01-31")
mdy("January 31st, 2017")
dmy("31-Jan-2017")
```

```{r}
ymd(20170131)
```

```{r}
ymd_hms("2017-01-31 20:11:59")
mdy_hm("01/31/2017 08:01")
```
```{r}
ymd(20170131, tz = "UTC")
```

```{r}
flights %>%
  select(year, month, day, hour, minute)
```

```{r}
flights %>%
  select(year, month, day, hour, minute) %>% 
  mutate(
    departure = make_datetime(year, month, day, hour, minute)
  )
```


```{r}
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}
flights_dt <- flights %>%
  filter(!is.na(dep_time), !is.na(arr_time)) %>%
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)) %>%
select(origin, dest, ends_with("delay"), ends_with("time"))

flights_dt
```

```{r}
flights_dt %>%
  ggplot(aes(dep_time)) + 
  geom_freqpoly(binwidth = 86400) # 86400 seconds = 1 day
```

```{r}
flights_dt %>%
  filter(dep_time < ymd(20130102)) %>% 
  ggplot(aes(dep_time)) + 
  geom_freqpoly(binwidth = 600)
```

```{r}
as_datetime(today())
as_date(now())
```

```{r}
as_datetime(60*60*10)
as_date(365*10+2)
```

```{r}
datetime <- ymd_hms("2016-07-08 12:34:56")

year(datetime)
month(datetime)
mday(datetime)
yday(datetime)
wday(datetime)
```

```{r}
month(datetime, label = TRUE)
wday(datetime, label = TRUE, abbr = FALSE)
```

```{r}
flights_dt %>%
  mutate(wday = wday(dep_time, label = TRUE)) %>%
  ggplot(aes(x = wday)) +
    geom_bar()
```

```{r}
flights_dt %>%
  mutate(minute = minute(dep_time)) %>%
  group_by(minute) %>%
  summarize(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  ) %>%
  ggplot(aes(minute, avg_delay)) + 
  geom_line()
```

```{r}
sched_dep <- flights_dt %>%
  mutate(minute = minute(sched_dep_time)) %>%
  group_by(minute) %>%
  summarize(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(sched_dep, aes(minute, avg_delay)) + 
  geom_line()
```

```{r}
ggplot(sched_dep, aes(minute, n)) + 
  geom_line()
```

```{r}
flights_dt %>%
  count(week = floor_date(dep_time, "week")) %>%
  ggplot(aes(week, n)) + 
  geom_line()
```

```{r}
(datetime <- ymd_hms("2016-07-08 12:34:56"))
```

```{r}
year(datetime) <- 2020
month(datetime) <- 01
hour(datetime) <- hour(datetime) + 1
datetime
```

```{r}
update(datetime, year = 2020, month = 2, mday = 2, hour = 2)
```

```{r}
ymd("2015-02-01") %>%
  update(mday = 30)
ymd("2015-02-01") %>%
  update(hour = 400)
```

```{r}
flights_dt %>%
  mutate(dep_hour = update(dep_time, yday = 1)) %>%
  ggplot(aes(dep_hour)) + 
  geom_freqpoly(binwidth = 300)
```

```{r}
h_age <- today() - ymd(19791014)
h_age
```

```{r}
as.duration(h_age)
```

```{r}
dseconds(15)
dminutes(10)
dhours(c(12,24))
ddays(0:5)
dweeks(3)
dyears(1)
```

```{r}
2 * dyears(1)
dyears(1) + dweeks(12) + dhours(15)
```

```{r}
tomorrow <- today() + ddays(1)
last_year <- today() - dyears(1)
```

```{r}
one_om <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")
one_pm
one_pm + ddays(1)
```

```{r}
one_pm
one_pm + days(1)
```

```{r}
seconds(15)
minutes(10)
hours(c(12,24))
days(7)
months(1:6)
weeks(3)
years(1)
```

```{r}
10 * (months(6) + days(1))
days(50) + hours(25) + minutes(2)
```

```{r}
ymd("2016-01-01") + dyears(1)
ymd("2016-01-01") + years(1)
```

```{r}
one_pm + ddays(1)
one_pm + days(1)
```

```{r}
flights_dt %>%
  filter(arr_time < dep_time)
```

```{r}
flights_dt <- flights_dt %>%
  mutate(overnight = arr_time < dep_time,
         arr_time = arr_time + days(overnight * 1),
         sched_arr_time = sched_arr_time + days(overnight * 1))
```

```{r}
flights_dt %>% 
  filter(overnight, arr_time < dep_time)
```

```{r}
years(1)/days(1)
```

```{r}
next_year <- today() + years(1)
(today() %--% next_year) / ddays(1)
```

```{r}
(today() %--% next_year) %/% days(1)
```

```{r}
Sys.timezone()
```

```{r}
length(OlsonNames())
head(OlsonNames())
```

```{r}
(x1 <- ymd_hms("2015-06-01 12:00:00", tz = "America/New_York"))
(x2 <- ymd_hms("2015-06-01 18:00:00", tz = "Europe/Copenhagen"))
(x3 <- ymd_hms("2015-06-02 04:00:00", tz = "Pacific/Auckland"))
```

```{r}
x1 - x2
x1 - x3
```

```{r}
x4 <- c(x1, x2, x3)
x4
```

```{r}
x4a <- with_tz(x4, tzone = "Australia/Lord_Howe")
x4a
x4a-x4
```

```{r}
x4b <- force_tz(x4, tzone = "Australia/Lord_Howe")
x4b
x4b-x4
```




# Part 3. Program 

## 14. Pipes with magrittr 

```{r}
library(magrittr)
```

```{r}
foo_foo <- little_bunny()
foo_foo_1 <- hop(foo_foo, through = forest)
foo_foo_2 <- scoop(foo_foo_1, up = field_mice)
foo_foo_3 <- bop(foo_foo_2, on = head)
```

```{r}
diamonds <- ggplot2::diamonds
diamonds2 <- diamonds %>%
  dplyr::mutate(price_per_carat = price / carat)

pryr::object_size(diamonds)
pryr::object_size(diamonds2)
pryr::object_size(diamonds, diamonds2)
```

```{r}
diamonds$carat[1] <- NA
pryr::object_size(diamonds)
pryr::object_size(diamonds2)
pryr::object_size(diamonds, diamonds2)
```

```{r}
foo_foo <- hop(foo_foo, through = forest)
foo_foo <- scoop(foo_foo, up = field_mice)
foo_foo <- bop(foo_foo, on = head)
```

```{r}
bop(
  scoop(
    hop(foo_foo, through = forest), 
    up = field_mice
  ),
  on = head
)
```

```{r}
foo_foo %>% 
  hop(through = forest) %>%
  scoop(up = field_mouse) %>%
  bop(on = head)
```

```{r}
my_pipe <- function(.) {
. <- hop(., through = forest)
. <- scoop(., up = field_mice)
bop(., on = head)
}
my_pipe(foo_foo)
```

```{r}
assign("x", 10)
x
"x" %>% assign(100)
x
```

```{r}
env <- environment()
"x" %>% assign(100, envir = env)
x
```

```{r}
tryCatch(stop("!"), error = function(e) "An error")
```

```{r}
stop("!") %>%
  tryCatch(error = function(e) "An error")
```

```{r}
rnorm(100) %>% 
  matrix(ncol = 2) %>%
  plot() %>%
  str()
```

```{r}
rnorm(100) %>%
  matrix(ncol = 2) %T>%
  plot() %>%
  str()
```

```{r}
mtcars %$%
  cor(disp, mpg)
```

```{r}
cor(mtcars$disp, mtcars$mpg)
```

```{r}
mtcars <- mtcars %>%
  transform(cyl = cyl * 2)
```

```{r}
mtcars %<>% transform(cyl = cyl * 2)
```


## 15. Functions 

```{r}
df <- tibble::tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

df$a <- (df$a - min(df$a, na.rm = TRUE)) / 
 (max(df$a, na.rm = TRUE) - min(df$a, na.rm = TRUE))
df$b <- (df$b - min(df$b, na.rm = TRUE)) /
(max(df$b, na.rm = TRUE) - min(df$a, na.rm = TRUE))
df$c <- (df$c - min(df$c, na.rm = TRUE)) /
(max(df$c, na.rm = TRUE) - min(df$c, na.rm = TRUE))
df$d <- (df$d - min(df$d, na.rm = TRUE)) /
(max(df$d, na.rm = TRUE) - min(df$d, na.rm = TRUE))
```

```{r}
x <- df$a
(x - min(x, na.rm = TRUE))/(max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
```

```{r}
rng <- range(x, na.rm = TRUE)
(x - rng[1]) / (rng[2] - rng[1])
```

```{r}
rescale01 <- function(x){
  rng <- range(x, na.rm=TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(c(0,5,10))
```

```{r}
rescale01(c(-10,0,10))
rescale01(c(1,2,3,NA,5))
```

```{r}
df$a <- rescale01(df$a)
df$b <- rescale01(df$b)
df$c <- rescale01(df$c)
df$d <- rescale01(df$d)
```

```{r}
x <- c(1:10, Inf)
rescale01(x)
```

```{r}
rescale01 <- function(x){
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1])/(rng[2]-rng[1])
}
rescale01(x)
```

```{r}
f()
my_awesome_function()
impute_missing()
collapse_years()
```

```{r}
col_mins <- function(x,y){}
rowMaxes <- function(x,y){}
```

```{r}
input_select()
input_checkbox()
imput_text()

select_input()
checkbox_input()
text_input()
```

```{r}
if(condition){
  # code executed when conditio is TRUE
} else{
  # code executed when condition is FALSE
}
```

```{r}
has_name <- function(x){
  nms <- name(x)
  if(is.null(nms)){
    rep(FALSE, length(x))
  } else{
    !is.na(nms) & nms!=""
  }
}
```

```{r}
if(c(TRUE, FALSE)){}
if(NA){}
```

```{r}
identical(0L,0)
x <- sqrt(2)^2
x
x==2
x-2
```

```{r}
if(this){
  # do that
} else if (that){
  # do something else
} else {
  #
}
```

```{r}
function(x,y,op){
  switch(op,
         plus = x + y,
         minus = x - y,
         times = x * y,
         divide = x / y,
         stop("Unknown op!"))
}
```

```{r}
if(y < 0 && debug){
  message("Y is negative")
}

if(y == 0){
  log(x)
} else{
  y ^ x
}
```

```{r}
y <- 10
x <- if(y<20) "Too low" else "Too high"
```

```{r}
if(y < 20){
  x <- "Too low"
} else{
  x <- "Too high"
}
```

```{r}
mean_ci <- function(x, conf = 0.95){
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - conf
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

x <- runif(100)
mean_ci(x)
mean_ci(x, conf = 0.99)
```

```{r}
mean(1:10, na.rm = TRUE)
```

```{r}
# x,y,z: vectors
# w: a vector of weights 
# df: a data frame 
# i,j: numeric indices (typically rows and columns)
# n: length, or number of rows
# p: number of columns 
```

```{r}
wt_mean <- function(x, w){
  sum(x * w) / sum(x)
}
wt_var <- function(x, w){
  mu <- wt_mean(x, w)
  sum(w * (x - mu) ^ 2) / sum(w)
}
wt_sd <- function(x, w){
  sqrt(wt_var(x, w))
}
```

```{r}
wt_mean(1:6, 1:3)
```

```{r}
wt_mean <- function(x, w){
  if(length(x) != length(w)){
    stop("`x` and `w` must be the same length", call = FALSE)
  }
  sum(w*x) / sum(x)
}
```

```{r}
wt_mean <- function(x, w, na.rm = FALSE){
  if(!is.logical(na.rm)){
    stop("`na.rm` must be logical")
  }
  if(length(na.rm) != 1){
    stop("`na.rm` must be length 1")
  }
  if(length(x) != length(w)){
    stop("`x` and `w` must be the same length", call.=FALSE)
  }
  if(na.rm){
    miss <- is.na(x) | is.na(w)
    x <- x[!miss]
    w <- w[!miss]
  }
  sum(w*x) / sum(x)
}
```

```{r}
wt_mean <- function(x, w, na.rm = FALSE){
  stopifnot(is.logical(na.rm), length(na.rm) == 1)
  stopifnot(length(x)==length(w))
  
  if(na.rm){
    miss <- is.na(x) | is.na(w)
    x <- x[!miss]
    w <- w[!miss]
  }
  sum(w * x) / sum(x)
}
wt_mean(1:6, 6:1, na.rm = "foo")
```

```{r}
sum(1,2,3,4,5,6,7,8,9,10)
```

```{r}
stringr::str_c("a","b","c","d","e","f")
```

```{r}
commas <- function(...) stringr::str_c(..., collapse = ", ")
commas(letters[1:10])
```

```{r}
rule <- function(..., pad = "-"){
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5
  cat(title, " ", stringr::str_dup(pad, width), "\n", sep="")
}
rule("Important output")
```

```{r}
x <- c(1,2)
sum(x, na.mr = TRUE)
```

```{r}
complicated_function <- function(x,y,z){
  if(length(x) == 0 || length(y) == 0){
    return(0)
  }
  # complicated code here
}
```

```{r}
f <- function(){
  if(x){
    # Do
    # something
    # that
    # takes 
    # many
    # lines 
    # to
    # express
  }else{
    # return something short 
  }
}
```

```{r}
f <- funtion(){
  if(!x){
    return(something_short)
  }
  # Do 
  # something
  # that
  # takes 
  # many 
  # lines 
  # to 
  # express
}
```

```{r}
show_missings <- function(df){
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
  invisible(df)
}
```

```{r}
show_missings(mtcars)
```

```{r}
x <- show_missings(mtcars)
class(x)
dim(x)
```

```{r}
mtcars %>% 
  show_missings() %>%
  mutate(mpg = ifelse(mpg < 20, NA, mpg)) %>%
  show_missings()
```

```{r}
f <- function(x){
  x + y
}
```

```{r}
y <- 100
f(10)
```

```{r}
y <- 1000
f(10)
```

```{r}
`+` <- function(x, y) {
if (runif(1) < 0.1) {
sum(x, y)
} else {
sum(x, y) * 1.1
}
}
table(replicate(1000, 1 + 2))
#>
#> 3 3.3
#> 100 900
rm(`+`)
```


## 16. Vectors 

```{r}
library(tidyverse)
```

```{r}
typeof(letters)
typeof(1:10)
x <- list("a","b",1:10)
length(x)
```

```{r}
1:10 %% 3 == 0
c(TRUE, TRUE, FALSE, NA)
```

```{r}
typeof(1)
typeof(1L)
1.5L
```

```{r}
x <- sqrt(2) ^ 2
x
x - 2
```

```{r}
c(-1, 0, 1) / 0
```

```{r}
is.finite(0)
is.infinite(Inf)
is.na(NA)
is.nan(NaN)
```

```{r}
x <- "This is a reasonably long string"
pryr::object_size(x)

y <- rep(x,1000)
pryr::object_size(y)
```

```{r}
NA
NA_integer_
NA_real_
NA_character_
```

```{r}
x <- sample(20, 100, replace = TRUE)
y <- x > 10
sum(y)
mean(y)
```

```{r}
typeof(c(TRUE, 1L))
typeof(c(1L, 1.5))
typeof(c(1.5, "a"))
```

```{r}
is_logical()
is_integer()
is_double()
is_numeric()
is_character()
is_atomic()
is_list()
is_vector()
```

```{r}
sample(10) + 100
runif(10) > 0.5
```

```{r}
1:10 + 1:2
```

```{r}
1:10 + 1:3
```

```{r}
tibble(x = 1:4,
       y = 1:2)
```

```{r}
tibble(x = 1:4,
       y = rep(1:2,2))
```

```{r}
tibble(x = 1:4,
       y = rep(1:2, each = 2))
```

```{r}
c(x = 1, y = 2, z = 4)
```

```{r}
set_names(1:3, c("a","b","c"))
```

```{r}
x <- c("one","two","three","four","five")
x[c(3,2,5)]
```

```{r}
x[c(1,1,5,5,2)]
```

```{r}
x[c(-1, -3, -5)]
```
```{r}
x[c(1,-1)]
```

```{r}
x[0]
```

```{r}
x <- c(10,3,NA,5,8,1,NA)
x[!is.na(x)]
x[x %% 2 == 0]
```

```{r}
x <- c(abc = 1, def = 2, xyz = 5)
x[c("xyz", "def")]
```

```{r}
x <- list(1,2,3)
x
```

```{r}
str(x)
```

```{r}
x_names <- list(a = 1, b = 2, c = 3)
str(x_named)
```

```{r}
y <- list("a",1L,1.5,TRUE)
str(y)
```

```{r}
z <- list(list(1,2),list(3,4))
str(z)
```

```{r}
x1 <- list(c(1,2),c(3,4))
x2 <- list(list(1,2),list(3,4))
x3 <- list(1,list(2,list(3)))
```

```{r}
a <- list(a = 1:3, b = "a string", c = pi, d = list(-1, -5))
```

```{r}
str(a[1:2])
```

```{r}
str(a[4])
```

```{r}
str(y[[1]])
```

```{r}
str(y[[4]])
```

```{r}
a$a
```

```{r}
a[["a"]]
```

```{r}
x <- 1:10
attr(x, "greeting")
```

```{r}
attr(x, "greeting")<-"Hi!"
attr(x, "farewell") <- "Bye!"
attributes(x)
```

```{r}
str(x)
```

```{r}
as.Date
```

```{r}
methods("as.Date")
```

```{r}
getS3method("as.Date","default")
```

```{r}
getS3method("as.Date","numeric")
```

```{r}
x <- factor(c("ab","cd","ab"), levels = c("ab","cd","ef"))
typeof(x)
attributes(x)
```

```{r}
x <- as.Date("1971-01-01")
unclass(x)
typeof(x)
attributes(x)
```

```{r}
x <- lubridate::ymd_hm("1970-01-01 01:00")
unclass(x)
typeof(x)
attributes(x)
```

```{r}
attr(x, "tzone") <- "US/Pacific"
x
attr(x, "tzone") <- "US/Eastern"
x
```

```{r}
y <- as.POSIXct(x)
typeof(y)
attributes(y)
```

```{r}
tb <- tibble::tibble(x = 1:5, y = 5:1)
typeof(tb)
attributes(tb)
```

```{r}
df <- data.frame(x = 1:5, y = 5:1)
typeof(df)
attributes(df)
```



## 17. Iteration with purrr


```{r}
library(tidyverse)
```

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
df
```

```{r}
median(df$a)
median(df$b)
median(df$c)
median(df$d)
```

```{r}
output <- vector("double",ncol(df)) # 1. output
for (i in seq_along(df)){           # 2. sequence
  output[[i]] <- median(df[[i]])    # 3. body
}
output
```

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
rescale01 <- function(x){
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

df$a <- rescale01(df$a)
df$b <- rescale01(df$b)
df$c <- rescale01(df$c)
df$d <- rescale01(df$d)
```

```{r}
for (i in seq_along(df)) {
  df[[i]] <- rescale01(df[[i]])
}
```

```{r}
for (i in seq_along(x)){
  name <- names(x)[[i]]
  value <- x[[i]]
}
```

```{r}
means <- c(0,1,2)

output <- double()
for (i in seq_along(means)){
  n <- sample(100,1)
  output <- c(output, rnorm(n, means[[i]]))
}
str(output)
```

```{r}
out <- vector("list", length(means))
for (i in seq_along(means)){
  n <- sample(100,1)
  out[[i]] <- rnorm(n,means[[i]])
}
str(out)
str(unlist(out))
```

```{r}
while(condition){
  # body
}
```

```{r}
for (i in seq_along(x)){
  # body
}

# Equivalent to
i <- 1
while(i <= length(x)){
  # body
  i <- i + 1
}
```

```{r}
flip <- function() sample(c("T","H"),1)

flips <- 0
nheads <- 0

while(nheads < 3) {
  if(flip() == "H"){
    nheads <- nheads + 1
  }else{
    nheads <- 0
  }
  flips <- flips + 1
}
flips
```

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

```{r}
output <- vector("double", length(df))
for (i in seq_along(df)) {
  output[[i]] <- mean(df[[i]])
}
output
```

```{r}
col_mean <- function(df){
  output <- vector("double",length(df))
  for (i in seq_along(df)){
    output[i] <- mean(df[[i]])
  }
  output
}
```

```{r}
col_median <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- median(df[[i]])
  }
  output
}

col_sd <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- sd(df[[i]])
  }
  output
}
```

```{r}
f1 <- function(x) abs(x - mean(x)) ^ 1
f2 <- function(x) abs(x - mean(x)) ^ 2
f3 <- function(x) abs(x - mean(x)) ^ 3
```

```{r}
f <- function(x, i) abs(x - mean(x)) ^ i
```

```{r}
col_summary <- function(df, fun){
  out <- vector("double", length(df))
  for(i in seq_along(df)){
    out[i] <- fun(df[[i]])
  }
  out
}

col_summary(df, median)
col_summary(df, mean)
```

```{r}
map()
map_lgl()
map_int()
map_dbl()
map_chr()
```

```{r}
map_dbl(df, mean)
map_dbl(df, median)
map_dbl(df, sd)
```

```{r}
df %>% map_dbl(mean)
df %>% map_dbl(median)
df %>% map_dbl(sd)
```

```{r}
map_dbl(df, mean, trim = 0.5)
```

```{r}
z <- list(x = 1:3, y = 4:5)
map_int(z, length)
```

```{r}
models <- mtcars %>%
  split(.$cyl) %>%
  map(function(df) lm(mpg~wt, data = df))
models
```

```{r}
models <- mtcars %>%
  split(.$cyl) %>%
  map(~lm(mpg~wt, data=.))
models
```

```{r}
models %>%
  map(summary) %>%
  map_dbl(~.$r.squared)
```

```{r}
models %>% 
  map(summary) %>%
  map_dbl("r.squared")
```

```{r}
x <- list(list(1,2,3), list(4,5,6), list(7,8,9))
x %>% map_dbl(2)
```

```{r}
x1 <- list(
c(0.27, 0.37, 0.57, 0.91, 0.20),
c(0.90, 0.94, 0.66, 0.63, 0.06),
c(0.21, 0.18, 0.69, 0.38, 0.77)
)
x2 <- list(
c(0.50, 0.72, 0.99, 0.38, 0.78),
c(0.93, 0.21, 0.65, 0.13, 0.27),
c(0.39, 0.01, 0.38, 0.87, 0.34)
)
threshold <- function(x, cutoff = 0.8) x[x > cutoff]
x1 %>% sapply(threshold) %>% str()
#> List of 3
#> $ : num 0.91
#> $ : num [1:2] 0.9 0.94
#> $ : num(0)
x2 %>% sapply(threshold) %>% str()
#> num [1:3] 0.99 0.93 0.87
```

```{r}
safe_log <- safely(log)
str(safe_log(10))
str(safe_log("a"))
```

```{r}
x <- list(1,10,"a")
y <- x %>% map(safely(log))
str(y)
```

```{r}
y <- y %>% transpose()
str(y)
```

```{r}
is_ok <- y$error %>% map_lgl(is_null)
x[!is_ok]
y$result[is_ok] %>% flatten_dbl
```

```{r}
x <- list(1,10,"a")
x %>% map_dbl(possibly(log, NA_real_))
```

```{r}
x <- list(1,-1)
x %>% map(quietly(log)) %>% str()
```

```{r}
mu <- list(5,10,-3)
mu %>% 
  map(rnorm, n = 5) %>%
  str()
```

```{r}
sigma <- list(1,5,10)
seq_along(mu) %>%
  map(~rnorm(5,mu[[.]],sigma[[.]])) %>%
  str()
```

```{r}
map2(mu, sigma, rnorm, n = 5) %>% str()
```

```{r}
mu %>% 
  map2(sigma, rnorm, n = 5) %>%
  str()
```

```{r}
map2 <- function(x,y,f,...){
  out <- vector("list",length(x))
  for(i in seq_along(x)){
    out[[i]]<-f(x[[i]],y[[i]],...)
  }
  out
}
```

```{r}
n <- list(1,3,5)
args1 <- list(n, mu, sigma)
args1 %>% 
  pmap(rnorm) %>%
  str()
```

```{r}
args2 <- list(mean = mu, sd = sigma, n = n)
args2 %>% 
  pmap(rnorm) %>%
  str()
```

```{r}
params <- tribble(
  ~mean, ~sd, ~n,
  5,1,1,
  10, 5, 3,
  -3, 10, 5
)
params %>%
  pmap(rnorm)
```

```{r}
args(rnorm)
```

```{r}
f <- c("runif","rnorm","rpois")
param <- list(
  list(min = -1, max = 1),
  list(sd = 5),
  list(lambda = 10)
)
```

```{r}
invoke_map(f,param,n=5) %>% str()
```

```{r}
sim <- tribble(
  ~f, ~params,
  "runif", list(min = 1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)
sim %>% mutate(sim = invoke_map(f,params,n=10))
```

```{r}
x <- list(1,"a",3)
x %>%
  walk(print)
```

```{r}
library(ggplot2)
plots <- mtcars %>%
  split(.$cyl) %>%
  map(~ggplot(.,aes(mpg,wt))+geom_point())
paths <- stringr::str_c(names(plots),".pdf")

pwalk(list(paths,plots),ggsave,path = tempdir())
```

```{r}
iris %>%
  keep(is.factor) %>%
  str()
```

```{r}
iris %>%
  discard(is.factor) %>% 
  str()
```

```{r}
x <- list(1:5, letters, list(10))
x %>%
  some(is_character)
```

```{r}
x %>% 
  every(is_vector)
```

```{r}
x <- sample(10)
x

x %>% 
  detect(~ .>5)

x %>%
  detect_index(~ .>5)
```

```{r}
x %>%
  head_while(~. >5)

x %>%
  tail_while(~. >5)
```

```{r}
dfs <- list(
  age = tibble(name = "John", age = 30),
  sex = tibble(name = c("John", "Mary"), sex = c("M","F")),
  trt = tibble(name = "Mary", treatment = "A")
)

dfs %>% reduce(full_join)
```

```{r}
vs <- list(
  c(1,3,5,6,10),
  c(1,2,3,7,8,10),
  c(1,2,3,4,8,9,10)
)
vs %>% reduce(intersect)
```

```{r}
x <- sample(10)
x
x %>% accumulate(`+`)
```

```{r}
x %>% reduce(`+`)
```




# Part 4. Model

## 18. Model basics with modelr 


```{r}
library(tidyverse)
library(modelr)
options(na.action = na.warn)
```

```{r}
ggplot(sim1, aes(x,y)) + geom_point()
```

```{r}
models <- tibble(
  a1 = runif(250, -20, 40),
  a2 = runif(250,-5,5)
)
ggplot(sim1, aes(x,y)) + 
  geom_abline(
    aes(intercept = a1, slope = a2),
    data = models, alpha = 1/4
  ) + 
  geom_point()
```

```{r}
model1 <- function(a, data){
  a[1] + data$x * a[2]
}
model1(c(7,1.5), sim1)
```

```{r}
measure_distance <- function(mod, data){
  diff <- data$y - model1(mod, data)
  sqrt(mean(diff^2))
}
measure_distance(c(7,1.5),sim1)
```

```{r}
sim1_dist <- function(a1,a2){
  measure_distance(c(a1,a2),sim1)
}
models <- models %>%
  mutate(dist = purrr::map2_dbl(a1,a2,sim1_dist))
models
```

```{r}
ggplot(sim1, aes(x,y))+
  geom_point(size = 2, color = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, color = -dist),
    data = filter(models, rank(dist)<=10)
  )
```

```{r}
ggplot(models, aes(a1,a2))+
  geom_point(
    data = filter(models, rank(dist) <= 10),
    size = 4, color = "red"
  )+ 
  geom_point(aes(colour = -dist))

```

```{r}
grid <- expand.grid(
  a1 = seq(-5,20,length = 25),
  a2 = seq(1,3,length = 25)
) %>%
  mutate(dist = purrr::map2_dbl(a1,a2,sim1_dist))

grid
```

```{r}
grid %>%
  ggplot(aes(a1,a2))+
  geom_point(
    data = filter(grid, rank(dist)<=10),
    size = 4, color = "red"
  ) + 
  geom_point(aes(color = -dist))
```

```{r}
ggplot(sim1, aes(x,y)) + 
  geom_point(size = 2, color = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, color = -dist),
    data = filter(grid, rank(dist)<=10)
  )
```

```{r}
best <- optim(c(0,0), measure_distance, data = sim1)
best$par

ggplot(sim1, aes(x,y)) + 
  geom_point(size = 2, color = "grey30") + 
  geom_abline(intercept = best$par[1], slope = best$par[2])
```

```{r}
sim1_mod <- lm(y~x, data = sim1)
coef(sim1_mod)
```

```{r}
grid <- sim1 %>%
  data_grid(x)
grid
```

```{r}
sim1
```

```{r}
grid <- grid %>%
  add_predictions(sim1_mod)
grid
```

```{r}
ggplot(sim1, aes(x)) + 
  geom_point(aes(y = y)) + 
  geom_line(
    aes(y = pred),
    data = grid,
    color = "red",
    size = 1
  )
```

```{r}
sim1 <- sim1 %>%
  add_residuals(sim1_mod)
sim1
```

```{r}
ggplot(sim1, aes(resid)) + 
  geom_freqpoly(binwidth = 0.5)
```

```{r}
ggplot(sim1, aes(x, resid)) + 
  geom_ref_line(h = 0) + 
  geom_point()
```

```{r}
df <- tribble(
  ~y, ~x1, ~x2,
  4,2,5,
  5,1,6
)
model_matrix(df, y ~ x1)
```

```{r}
model_matrix(df, y ~ x1 - 1)
```

```{r}
model_matrix(df, y ~ x1 + x2)
```

```{r}
df <- tribble(
  ~sex, ~response,
  "male",1,
  "female",2,
  "male",1
)
model_matrix(df, response ~ sex)
```

```{r}
ggplot(sim2) + 
  geom_point(aes(x,y))
```

```{r}
mod2 <- lm(y ~ x, data = sim2)
grid <- sim2 %>%
  data_grid(x) %>%
  add_predictions(mod2)
grid
```

```{r}
ggplot(sim2, aes(x)) + 
  geom_point(aes(y = y)) + 
  geom_point(
    data = grid,
    aes(y = pred),
    color = "red",
    size = 4
  )
```

```{r}
tibble(x = "e") %>%
  add_predictions(mod2)
```

```{r}
ggplot(sim3, aes(x1,y)) + 
  geom_point(aes(color = x2))
```

```{r}
mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)
```

```{r}
grid <- sim3 %>%
  data_grid(x1,x2) %>%
  gather_predictions(mod1,mod2)
grid
```

```{r}
ggplot(sim3, aes(x1,y,color = x2)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) + 
  facet_wrap(~model)
```

```{r}
sim3 <- sim3 %>%
  gather_residuals(mod1, mod2)

ggplot(sim3, aes(x1, resid, color = x2)) + 
  geom_point() + 
  facet_grid(model ~ x2)
```

```{r}
mod1 <- lm(y ~ x1 + x2, data = sim4)
mod2 <- lm(y ~ x1 * x2, data = sim4)

grid <- sim4 %>%
  data_grid(
    x1 = seq_range(x1,5),
    x2 = seq_range(x2,5)
  ) %>%
  gather_predictions(mod1, mod2)
grid
```

```{r}
seq_range(c(0.0123,0.923423),n=5)
seq_range(c(0.0123,0.923423),n=5,pretty=TRUE)
```

```{r}
x1 <- rcauchy(100)
seq_range(x1,n=5)
seq_range(x1,n=5,trim=0.1)
seq_range(x1,n=5,trim=0.25)
seq_range(x1,n=5,trim=0.5)
```

```{r}
x2 <- c(0,1)
seq_range(x2, n = 5)
seq_range(x2,n=5,expand=0.1)
seq_range(x2,n=5,expand=0.25)
seq_range(x2,n=5,expand=0.5)
```

```{r}
ggplot(grid, aes(x1,x2)) + 
  geom_tile(aes(fill = pred)) + 
  facet_wrap(~model)
```

```{r}
ggplot(grid, aes(x1,pred,color = x2,group = x2)) + 
  geom_line() + 
  facet_wrap(~model)
```

```{r}
ggplot(grid, aes(x2, pred, color = x1, group = x1)) + 
  geom_line() + 
  facet_wrap(~model)
```

```{r}
df <- tribble(
  ~y, ~x,
  1,1,
  2,2,
  3,3
)
model_matrix(df, y ~ x^2 + x)
```

```{r}
model_matrix(df, y ~ I(x^2) + x)
```

```{r}
model_matrix(df, y ~ poly(x,2))
```
```{r}
library(splines)
model_matrix(df, y ~ ns(x,2))
```

```{r}
sim5 <- tibble(
  x = seq(0, 3.5 * pi, length = 50),
  y = 4 * sin(x) + rnorm(length(x))
)
ggplot(sim5, aes(x,y)) + 
  geom_point()
```

```{r}
mod1 <- lm(y ~ ns(x, 1), data = sim5)
mod2 <- lm(y ~ ns(x, 2), data = sim5)
mod3 <- lm(y ~ ns(x, 3), data = sim5)
mod4 <- lm(y ~ ns(x, 4), data = sim5)
mod5 <- lm(y ~ ns(x, 5), data = sim5)
```

```{r}
grid <- sim5 %>%
  data_grid(x = seq_range(x,n=50,expand=0.1)) %>%
  gather_predictions(mod1,mod2,mod3,mod4,mod5,.pred="y")

ggplot(sim5, aes(x,y)) + 
  geom_point() + 
  geom_line(data = grid, color = "red") + 
  facet_wrap(~model)
```

```{r}
df <- tribble(
  ~x, ~y,
  1,2.2,
  2,NA,
  3,3.5,
  4,8.3,
  NA,10
)
mod <- lm(y ~ x, data = df)
```

```{r}
mod <- lm(y ~ x, data = df, na.action = na.exclude)
nobs(mod)
```

```{r}
# generalized linear models 
# generalized addictive models 
# penalized linear models 
# robust linear models 
# trees 
```



## 19. Model building 

```{r}
library(tidyverse)
library(modelr)
options(na.action = na.warn)
library(nycflights13)
library(lubridate)
```

```{r}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
ggplot(diamonds, aes(color, price)) + geom_boxplot()
ggplot(diamonds, aes(clarity, price)) + geom_boxplot()
```

```{r}
ggplot(diamonds, aes(carat, price)) + geom_hex(bins = 50)
```

```{r}
diamonds2 <- diamonds %>%
  filter(carat <= 2.5) %>%
  mutate(lprice = log2(price), lcarat = log2(carat))
```

```{r}
ggplot(diamonds2, aes(lcarat, lprice)) + geom_hex(bins = 50)
```

```{r}
mod_diamond <- lm(lprice ~ lcarat, data = diamonds2)
```

```{r}
grid <- diamonds2 %>%
  data_grid(carat = seq_range(carat,20)) %>%
  mutate(lcarat = log2(carat)) %>%
  add_predictions(mod_diamond,"lprice") %>%
  mutate(price = 2 ^ lprice)

ggplot(diamonds2, aes(carat, price)) + 
  geom_hex(bins = 50) + 
  geom_line(data = grid, color = "red", size = 1)
```

```{r}
diamonds2 <- diamonds2 %>%
  add_residuals(mod_diamond, "lresid")
ggplot(diamonds2, aes(lcarat, lresid)) + 
  geom_hex(bins = 50)
```

```{r}
ggplot(diamonds2, aes(cut, lresid)) + geom_boxplot()
ggplot(diamonds2, aes(color, lresid)) + geom_boxplot()
ggplot(diamonds2, aes(clarity, lresid)) + geom_boxplot()
```

```{r}
mod_diamond2 <- lm(
  lprice ~ lcarat + color + cut + clarity,
  data = diamonds2
)
```

```{r}
grid <- diamonds2 %>%
  data_grid(cut, .model = mod_diamond2) %>%
  add_predictions(mod_diamond2)
grid
```

```{r}
ggplot(grid, aes(cut, pred)) + 
  geom_point()
```

```{r}
diamonds2 <- diamonds2 %>%
  add_residuals(mod_diamond2,"lresid2")
ggplot(diamonds2, aes(lcarat, lresid2)) + 
  geom_hex(bins = 50)
```

```{r}
diamonds2 %>%
  filter(abs(lresid2) > 1) %>%
  add_predictions(mod_diamond2) %>%
  mutate(pred = round(2^pred)) %>%
  select(price, pred, carat:table, x:z) %>%
  arrange(price)
```

```{r}
daily <- flights %>%
  mutate(date = make_date(year, month, day)) %>%
  group_by(date) %>%
  summarize(n = n())
daily
```

```{r}
ggplot(daily, aes(date, n)) + geom_line()
```

```{r}
daily <- daily %>%
  mutate(wday = wday(date, label = TRUE))
ggplot(daily, aes(wday, n)) + 
  geom_boxplot()
```

```{r}
mod <- lm(n ~ wday, data = daily)

grid <- daily %>% 
  data_grid(wday) %>%
  add_predictions(mod, "n")

ggplot(daily, aes(wday, n)) + 
  geom_boxplot() + 
  geom_point(data = grid, color = "red", size = 4)
```

```{r}
daily <- daily %>%
  add_residuals(mod)
daily %>%
  ggplot(aes(date, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line()
  
```

```{r}
ggplot(daily, aes(date, resid, color = wday)) + 
  geom_ref_line(h = 0) + 
  geom_line()
```

```{r}
daily %>%
  filter(resid < -100)
```

```{r}
daily %>%
  ggplot(aes(date, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line(color = "grey50") + 
  geom_smooth(se = FALSE, span = 0.2)
```

```{r}
daily %>%
  filter(wday == "Sat") %>%
  ggplot(aes(date, n)) + 
  geom_point() + 
  geom_line() + 
  scale_x_date(
    NULL,
    date_breaks = "1 month",
    date_labels = "%b"
  )
```

```{r}
term <- function(date){
  cut(date,
      breaks = ymd(20130101, 20130605, 20130825, 20140101),
      labels = c("spring","summer","fall"))
}

daily <- daily %>%
  mutate(term = term(date))

daily %>%
  filter(wday == "Sat") %>%
  ggplot(aes(date, n, color = term)) + 
  geom_point(alpha = 1/3) + 
  geom_line() + 
  scale_x_date(
    NULL,
    date_breaks = "1 month",
    date_labels = "%b"
  )
```

```{r}
daily %>%
  ggplot(aes(wday, n, color = term)) + 
  geom_boxplot()
```

```{r}
mod1 <- lm(n ~ wday, data = daily)
mod2 <- lm(n ~ wday * term, data = daily)

daily %>%
  gather_residuals(without_term = mod1, with_term = mod2) %>%
  ggplot(aes(date, resid, color = model)) + 
  geom_line(alpha = 0.75)
```

```{r}
grid <- daily %>%
  data_grid(wday, term) %>%
  add_predictions(mod2, "n")

ggplot(daily, aes(wday, n)) + 
  geom_boxplot() + 
  geom_point(data = grid, color = "red") + 
  facet_wrap(~term)
```

```{r}
mod3 <- MASS::rlm(n ~ wday * term, data = daily)

daily %>%
  add_residuals(mod3, "resid") %>%
  ggplot(aes(date, resid)) + 
  geom_hline(yintercept = 0, size = 2, color = "white") + 
  geom_line()
```

```{r}
compute_vars <- function(data){
  data %>%
    mutate(
      term = term(date),
      wday = wday(date, label = TRUE)
    )
}
```

```{r}
wday2 <- function(x) wday(x, label = TRUE)
mod3 <- lm(n ~ wday2(date) * term(date), data = daily)
```

```{r}
library(splines)
mod <- MASS::rlm(n ~ wday * ns(date,5), data = daily)

daily %>%
  data_grid(wday, date = seq_range(date, n = 13)) %>%
  add_predictions(mod) %>%
  ggplot(aes(date, pred, color = wday)) + 
  geom_line() + 
  geom_point()
```



## 20. Many models with purrr and broom

```{r}
library(modelr)
library(tidyverse)
```

```{r}
library(gapminder)
gapminder
```

```{r}
gapminder %>%
  ggplot(aes(year, lifeExp, group = country)) + 
  geom_line(alpha = 1/3)
```

```{r}
nz <- filter(gapminder, country == "New Zealand")
nz %>%
  ggplot(aes(year, lifeExp)) + 
  geom_line() + 
  ggtitle("Full data = ")
```

```{r}
nz_mod <- lm(lifeExp ~ year, data = nz)
nz %>%
  add_predictions(nz_mod) %>%
  ggplot(aes(year, pred)) + 
  geom_line() + 
  ggtitle("Linear trend + ")
```

```{r}
nz %>%
  add_residuals(nz_mod) %>%
  ggplot(aes(year, resid)) + 
  geom_hline(yintercept = 0, color = "white", size = 3) + 
  geom_line() + 
  ggtitle("Remaining pattern")
```

```{r}
by_country <- gapminder %>%
  group_by(country, continent) %>%
  nest()

by_country
```

```{r}
by_country$data[[1]]
```

```{r}
country_model <- function(df){
  lm(lifeExp ~ year, data = df)
}
```

```{r}
models <- map(by_country$data, country_model)
```

```{r}
by_country <- by_country %>%
  mutate(model = map(data, country_model))
by_country
```

```{r}
by_country %>%
  filter(continent == "Europe")
```

```{r}
by_country %>%
  arrange(continent, country)
```

```{r}
by_country <- by_country %>%
  mutate(
    resids = map2(data, model, add_residuals)
  )
by_country
```

```{r}
by_country$resids[[1]]
```

```{r}
resids <- unnest(by_country, resids)
resids
```

```{r}
resids %>%
  ggplot(aes(year, resid)) + 
  geom_line(aes(group = country), alpha = 1/3) + 
  geom_smooth(se = TRUE)
```

```{r}
resids %>%
  ggplot(aes(year, resid, group = country)) + 
  geom_line(alpha = 1/3) + 
  facet_wrap(~continent)
```

```{r}
broom::glance(nz_mod)
```

```{r}
by_country %>%
  mutate(glance = map(model, broom::glance)) %>%
  unnest(glance)
```

```{r}
glance <- by_country %>%
  mutate(glance = map(model, broom::glance)) %>%
  unnest(glance, .drop = TRUE)
glance
```

```{r}
glance %>%
  arrange(r.squared)
```

```{r}
glance %>%
  ggplot(aes(continent, r.squared)) + 
  geom_jitter(width = 0.5)
```

```{r}
bad_fit <- filter(glance, r.squared < 0.25)

gapminder %>%
  semi_join(bad_fit, by = "country") %>%
  ggplot(aes(year, lifeExp, color = country)) + 
  geom_line()
```

```{r}
data.frame(x = list(1:3, 3:5))
```

```{r}
data.frame(
  x = I(list(1:3, 3:5)),
  y = c("1,2","3,4,5")
)
```

```{r}
tibble(
  x = list(1:3, 3:5),
  y = c("1,2", "3,4,5")
)
```

```{r}
tribble(
  ~x, ~y,
  1:3, "1,2",
  3:5, "3,4,5"
)
```

```{r}
gapminder %>%
  group_by(country, continent) %>%
  nest()
```

```{r}
gapminder %>%
  nest(year:gdpPercap)
```

```{r}
df <- tribble(
  ~x1,
  "a,b,c",
  "d,e,f,g"
)
df
```

```{r}
df %>%
  mutate(x2 = stringr::str_split(x1, ","))
```

```{r}
df %>%
  mutate(x2 = stringr::str_split(x1, ",")) %>%
  unnest()
```

```{r}
sim <- tribble(
  ~f, ~params,
  "runif",list(min = -1, max = -1),
  "rnorm",list(sd = 5),
  "rpois",list(lambda = 10)
)

sim %>%
  mutate(sims = invoke_map(f, params, n = 10))
```

```{r}
mtcars %>%
  group_by(cyl) %>%
  summarize(q = quantile(mpg))
```

```{r}
mtcars %>%
  group_by(cyl) %>% 
  summarize(q = list(quantile(mpg)))
```

```{r}
probs <- c(0.01, 0.25, 0.5, 0.75, 0.99)
mtcars %>%
  group_by(cyl) %>%
  summarize(p = list(probs), q = list(quantile(mpg, probs))) %>%
  unnest()
```

```{r}
x <- list(
  a = 1:5,
  b = 3:4,
  c = 5:6
)
```

```{r}
df <- enframe(x)
df
```

```{r}
df %>%
  mutate(
    smry = map2_chr(
      name,
      value, ~ stringr::str_c(.x, ":", .y[1])
    )
  )
```

```{r}
df <- tribble(
  ~x,
  letters[1:5],
  1:3,
  runif(5)
)
df
```

```{r}
df %>%
  mutate(
    type = map_chr(x, typeof),
    length = map_int(x, length)
  )
```

```{r}
df <- tribble(
  ~x,
  list(a = 1, b = 2),
  list(a = 2, c = 4)
)
df
```

```{r}
df %>%
  mutate(
    a = map_dbl(x, "a"),
    b = map_dbl(x, "b", .null = NA_real_)
  )
```

```{r}
tibble(x = 1:2, y = list(1:4, 1)) %>% unnest(y)
```

```{r}
df1 <- tribble(
  ~x, ~y, ~z,
  1, c("a","b"), 1:2,
  2, "c", 3
)
df1
```

```{r}
df1 %>% unnest(y,z)
```

```{r}
df2 <- tribble(
  ~x, ~y, ~z,
  1, "a", 1:2,
  2, c("b","c"),3
)
df2
```

```{r}
df2 %>% unnest(y,z)
```



# Part 5. Communicate 

##21. R markdown 

```{r,
eval = FALSE,
include = FALSE,
echo = FALSE,
message = FALSE,
warning = FALSE,
results = 'hide',
fig.show = 'hide'
error = TRUE
}
```
```{r}
mtcars[1:5, 1:10]
```

```{r}
knitr::kable(
  mtcars[1:5,],
  caption = "A knitr kable."
)
```

```{r, cache = TRUE, dependson = ""}
knitr::clean_cache()
```

```{r}
# knitr::opts_chunk$set(
#   comment = "#>",
#   collapse = TRUE)

# knitr::opts_chunk$set(
# echo = FALSE
# )
```


```{r}
# we have about `r nrow(diaamonds)` diamonds. Only `r nrow(diamonds) - nrow(smaller)` are larger than 2.5 carats. The distribution of the remainder is shown below:
```


```{r}
comma <- function(x) format(x, digits = 2, big.mark = ",")
comma(3452345)
comma(.12358124331)
```



```{r}
# --- 
# output: html_document
# params: my_class:"suv"
---
# ```{r setup, include = FALSE}
# library(ggplot2)
# library(dplyr)
# class <- mpg %>% filter(class == params$my_class)
# ```
# # Fuel economy for `r params$my_class`s
# ```{r, message = FALSE}
# ggplot(class, aes(displ, hwy)) +
# geom_point() +
# geom_smooth(se = FALSE)
# ```
```

```{r}
# params:
# start: !r lubridate::ymd("2015-01-01")
# snapshot: !r lubridate::ymd_hms("2015-01-01 12:30:00")
```

```{r}
# rmarkdown::render(
#   "fuel-economy.Rmd",
#   params = list(my_class = "suv")
# )
```

```{r}
# reports <- tibble(
#  class = unique(mpg$class),
#  filename = stringr::str_c("fuel-economy-",class,".html"),
#  params = purrr:map(class, ~list(my_class = .))
#)
#reports
```



```{r}
#reports %>%
#  select(output_file = filename, params) %>%
#  purrr::pwalk(rmarkdown::render, input = "fuel-economy.Rmd")
```





## 22. Graphics for communication with ggplot2 

```{r}
library(tidyverse)
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  geom_smooth(se = TRUE) + 
  labs(
    title = paste(
      "Fuel efficiency generally decreases with engine size"
    )
  )
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
geom_point(aes(color = class)) +
geom_smooth(se = FALSE) +
labs(
title = paste(
"Fuel efficiency generally decreases with", "engine size"),
subtitle = paste(
"Two seaters (sports cars) are an exception", "because of their light weight"),
caption = "Data from fueleconomy.gov"
)
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  geom_smooth(se = FALSE) + 
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    color = "Car type"
  )
```

```{r}
df <- tibble(
x = runif(10),
y = runif(10)
)
ggplot(df, aes(x, y)) +
geom_point() +
labs(
x = quote(sum(x[i] ^ 2, i == 1, n)),
y = quote(alpha + beta + frac(delta, theta))
)
```


```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  geom_text(aes(label = model), data = best_in_class)
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_label(
    aes(label = model),
    data = best_in_class,
    nudge_y = 2,
    alpha = 0.5
)
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  geom_point(size = 3, shape = 1, data = best_in_class) + 
  ggrepel::geom_label_repel(
    aes(label = model),
    data = best_in_class
  )
```

```{r}
class_avg <- mpg %>%
group_by(class) %>%
summarize(
displ = median(displ),
hwy = median(hwy)
)

ggplot(mpg, aes(displ, hwy, color = class)) +
ggrepel::geom_label_repel(aes(label = class),
data = class_avg,
size = 6,
label.size = 0,
segment.color = NA
) +
geom_point() +
theme(legend.position = "none")
```

```{r}
label <- mpg %>%
summarize(
displ = max(displ),
hwy = max(hwy),
label = paste(
"Increasing engine size is \nrelated to decreasing fuel economy."
)
)
ggplot(mpg, aes(displ, hwy)) +
geom_point() +
geom_text(
aes(label = label),
data = label,
vjust = "top",
hjust = "right"
)
```

```{r}
label <- tibble(
displ = Inf,
hwy = Inf,
label = paste(
"Increasing engine size is \nrelated to decreasing fuel economy."
)
)
ggplot(mpg, aes(displ, hwy)) +
geom_point() +
geom_text(
aes(label = label),
data = label,
vjust = "top",
hjust = "right"
)
```

```{r}
"Increasing engine size related to decreasing fuel economy." %>%
stringr::str_wrap(width = 40) %>%
writeLines()
```

```{r}
# geom_hline(), geom_vline(), geom_rect(), geom_segment() 
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class))
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete()
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_y_continuous(breaks = seq(15,40,by = 5))
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_x_date(
      NULL,
      breaks = presidential$start,
      date_labels = "'%y"
)
```

```{r}
base <- ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right")
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  guides(
    color = guide_legend(
    nrow = 1,
    override.aes = list(size = 4)
  )
)
```

```{r}
ggplot(diamonds, aes(carat, price)) + 
  geom_bin2d()

ggplot(diamonds, aes(log10(carat), log10(price))) + 
  geom_bin2d()
```

```{r}
ggplot(diamonds, aes(carat, price)) + 
  geom_bin2d() + 
  scale_x_log10() + 
  scale_y_log10()
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
geom_point(aes(color = drv)) +
scale_color_brewer(palette = "Set1")
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_color_brewer(palette = "Set1")
```

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, color = party)) + 
  geom_point() + 
  geom_segment(aes(xend = end, yend = id)) + 
  scale_color_manual(
    values = c(Republican = "red", Democratic = "blue")
  )
```

```{r}
df <- tibble(
x = rnorm(10000),
y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
geom_hex() +
coord_fixed()
```

```{r}
ggplot(df, aes(x, y)) +
geom_hex() +
viridis::scale_fill_viridis() +
coord_fixed()
```

```{r}
ggplot(mpg, mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) + 
  geom_smooth()
```

```{r}
ggplot(mpg, mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) + 
  geom_smooth() + 
  coord_cartesian(xlim = c(5,7), ylim = c(10,30))
```

```{r}
mpg %>%
  filter(displ >= 5, displ <= 7, hwy >= 10, hwy <= 30) %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()
```

```{r}
suv <- mpg %>% filter(class == "suv")
compact <- mpg %>% filter(class == "compact")

ggplot(suv, aes(displ, hwy, color = drv)) + geom_point()
```

```{r}
ggplot(compact, aes(displ, hwy, color = drv)) + geom_point()
```

```{r}
x_scale <- scale_x_continuous(limits = range(mpg$displ))
y_scale <- scale_y_continuous(limits = range(mpg$hwy))
col_scale <- scale_color_discrete(limits = unique(mpg$drv))

ggplot(suv, aes(displ, hwy, color = drv)) + 
  geom_point() + 
  x_scale + 
  y_scale + 
  col_scale
```

```{r}
ggplot(compact, aes(displ, hwy, color = drv)) + 
  geom_point() + 
  x_scale + 
  y_scale + 
  col_scale
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  geom_smooth() + 
  theme_bw()
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) + 
  geom_smooth() + 
  theme_dark()
```

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point()
```

```{r}
# ggsave("my_plot.pdf)
```

```{r}
# fig.width, fig.height, fig.asp, out.width, out.height
```



## 23. R markdown formats 

```{r}
# --- 
# title: "viridis Demo"
# output: html_document
# ---
```


```{r}
# rmarkdown::render(
#  "diamond-sizes.Rmd",
#  output_format = "word_document"
# )
```

```{r}
# flexdashboard
# htmlwidgets
```

```{r}
library(leaflet)
leaflet() %>%
setView(174.764, -36.877, zoom = 16) %>%
addTiles() %>%
addMarkers(174.764, -36.877, popup = "Maungawhau")
```

```{r}
library(shiny)
textInput("name", "What is your name?")
numericInput("age", "How old are you?", NA, min = 0, max = 150)
```


## 24. R markdown workflow 

```{r}
#
```

